---
title: "Practical Machine Learning"
author: "Marissa Arager"
date: "9/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, message=FALSE, warning=FALSE}
#Load Packages
library(dplyr)
library(caret)
library(AppliedPredictiveModeling)
library(data.table)
library(rattle)
library(ggplot2)
library(randomForest)

```


#### Background
The goal of this project was select a machine learning alogrithim to predict the manner in which test participants conducted exercises using the database from the following URL. http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. The data comes from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and was used to quantify how well a particluar exercie activity has been completed.


#### Data Cleaning
```{r, message=FALSE, warning=FALSE}

#load data and data cleaning 
data<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),na.strings=c("","NA"))
set.seed(1000)
# Remove unnecessary data columns
data<-data[,-c(1:7)]
# Remove columns that have more than 50% of missing data 
data<-data[,colSums(is.na(data))<.5*19622]
```

#### Model Setup
```{r, message=FALSE, warning=FALSE}
#Model Setup
#Split the data into training and testing sets
intrain<-createDataPartition(data$classe, p = .9, list = FALSE)
training<-data[intrain,]
testing<-data[-intrain,]
# Split the training set into subtraining and validation sets for use after final model has been selected 
insubtrain<-createDataPartition(training$classe, p = .8, list = FALSE)
subtraining<-training[insubtrain,]
validation<-training[-insubtrain,]
```

####Model Creation
```{r, message=FALSE, warning=FALSE}
#Model Creating 
#decision tree model
DTmodFit <- train(classe ~.,method="rpart",data=subtraining)
DTPred <- predict(DTmodFit, validation)
#Confusion Matrix
confusionMatrix(validation$classe, DTPred)

#Random Forest
RFmodFit<-randomForest(formula = classe ~ ., data = subtraining)
RFPred<-predict(RFmodFit, validation)
confusionMatrix(RFPred, validation$classe)
```

#### Model Selection
```{r, message=FALSE, warning=FALSE}
##predict final test set -- choose model
# By comparing the confusion matricies between the two models, Random forest is more accurate than the decision tree, so we will select Random Forest as our final model. 
FinalModPred<-predict(RFmodFit,testing)
confusionMatrix(FinalModPred, testing$classe)

```

#### Conclusion
By using a random forest algorithim we were able to predict how well a particular activity was conducted. 
